Computing exact restricted isometry constants via\\ mixed integer semidefinite programming

### 2017-03-06 RIP-MISDP-Paper/IEEE_RIP-MISDP_peerreview.tex
PAPER: no longer use the term "diagonal dominance"
PAPER: number I, II, III instead of A,B,C in table 1 since A,B,C are used differently in table 4
merged master

Our computational results show that the resulting problems are quite
hard to solve, even for matrices of smaller size. However, this is true for
the SDP relaxations of d'Aspremont et al.\ as well, and it allows for an
evaluation of the quality of these bounds.  In fact, these bounds turn out
to be quite tight or weak, depending on the instance. This complements the
theoretical approximation guarantees developed by d'Aspremont et
al. in~\cite{AspBG14}.

@@

  On the other hand, we can lift any solution $\check{X}$ of~\eqref{Proj}
  to a point $\hat{X} \in S_n$ by extending with zeros. Defining
  $\hat{z}_i$ to be 1 if and only if $i \in T$, yields a feasible point
  $(\hat{X}, \hat{z})$ for \eqref{MISDP} with identical objective value,
  since $\Tr(\check{X}) = 1$ and $\check{X} \in S_\ell$ imply $0 \leq
  \check{X}_{jj} = \hat{X}_{i_j i_j} \leq 1$ for $j = 1, \dots,
  \ell$. Therefore, $\hat{X}_{ij} \leq \hat{z}_j$ by diagonal dominance of
  semidefinite matrices.

@@

\begin{table} 
 \begin{scriptsize} \caption{Sizes of the used matrices; three matrices were
 randomly generated for each size} 
 \label{matrixsizes} 
 \begin{tabular*}{\linewidth}{@{}l@{\;\;\extracolsep{\fill}}rrrrrrrrr@{}}\toprule 
  & \multicolumn{3}{c}{A} & \multicolumn{3}{c}{B} & \multicolumn{3}{c}{C} \\ 
\cmidrule(r){2-4} \cmidrule{5-7}  \cmidrule(l){8-10}
type & $m$ & $n$ & $k$ & $m$ & $n$ & $k$ & $m$ & $n$ & $k$  \\ 
\midrule 
$N(0,1)$ & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
binary & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
band matrix & 30 & 30 & 5 & 35 & 35 & 4 & 40 & 40 & 3 \\
rank 1& 30 & 30 & 5 & 35 & 35 & 4 & 40 & 40 & 3 \\
$N(0,1/m)$ & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
$\pm 1/\sqrt{m}$ & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
$0, \pm \sqrt{3/m}$ & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
\bottomrule 
 \end{tabular*} 
 \end{scriptsize} 
 \end{table}

>>>>

Our computational results show that the resulting problems are quite
hard to solve, even for matrices of smaller size. However, they are
still solvable for larger instances than a brute-force approach computing
the eigenvalues of all submatrices of corresponding size.

Even if the techniques for solving MISDPs are not yet capable of solving
large enough instances to be used for computing RICs in practical applications,
by computing exact restricted isometry constants, our approach allows to evaluate
relaxations like those proposed by d'Aspremont et al. in \cite{Asp08,Asp07}.
In fact, these bounds turn out
to be quite tight or weak, depending on the instance, while in parts requiring
solution times comparable to those of the MISDP-approach. This complements the
theoretical approximation guarantees developed by d'Aspremont et
al. in~\cite{AspBG14}.

@@

  On the other hand, we can lift any solution $\check{X}$ of~\eqref{Proj}
  to a point $\hat{X} \in S_n$ by extending with zeros. Defining
  $\hat{z}_i$ to be 1 if and only if $i \in T$, yields a feasible point
  $(\hat{X}, \hat{z})$ for \eqref{MISDP} with identical objective value,
  since $\Tr(\check{X}) = 1$ and $\check{X} \in S_\ell$ imply $0 \leq
  \check{X}_{jj} = \hat{X}_{i_j i_j} \leq 1$ for $j = 1, \dots,
  \ell$. Therefore, $\hat{X}_{ij} \leq \sqrt{\hat{X}_{ii}\hat{X}_{jj}} \leq \hat{z}_j$ by 
  nonnegativity of the minors of positive semidefinite matrices.

@@

\begin{table} 
 \begin{scriptsize} \caption{Sizes of the used matrices; three matrices were
 randomly generated for each size} 
 \label{matrixsizes} 
 \begin{tabular*}{\linewidth}{@{}l@{\;\;\extracolsep{\fill}}rrrrrrrrr@{}}\toprule 
  & \multicolumn{3}{c}{I} & \multicolumn{3}{c}{II} & \multicolumn{3}{c}{III} \\ 
\cmidrule(r){2-4} \cmidrule{5-7}  \cmidrule(l){8-10}
type & $m$ & $n$ & $k$ & $m$ & $n$ & $k$ & $m$ & $n$ & $k$  \\ 
\midrule 
$N(0,1)$ & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
binary & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
band matrix & 30 & 30 & 5 & 35 & 35 & 4 & 40 & 40 & 3 \\
rank 1& 30 & 30 & 5 & 35 & 35 & 4 & 40 & 40 & 3 \\
$N(0,1/m)$ & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
$\pm 1/\sqrt{m}$ & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
$0, \pm \sqrt{3/m}$ & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
\bottomrule 
 \end{tabular*} 
 \end{scriptsize} 
 \end{table}

### 2017-03-06 RIP-MISDP-Paper/Paper_v1.tex
merged master

Now let $\bar{X}$ be the matrix consisting of those rows and columns 
$i_k$ of $\hat{X}$ such that $\hat{z}_{i_k} = 1$. Then $\bar{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), as we only removed zero rows and columns of $\hat{X}$ 
and $(A\T A)\hat{X}$, which do not influence the trace or the positive
semidefiniteness.

On the other hand we can lift any optimal solution $\check{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $\bar{z}_i$ to one if and only if $i \in T$
and $X^{\prime}_{i_j i_k} = \check{X}_{jk}$ if and only if $i_j, i_k \in T$
and $X^{\prime}_{ij}=0$ otherwise. This solution will then be feasible for (\ref{MISDP}), as $\Tr(\check{X}) = 1$ and $\check{X}\in S_\ell$ imply $0 \leq \check{X}_{jj} = X^{\prime}_{i_j i_j} \leq 1$ and therefore 
$X^{\prime}_{ij} \leq z^{\prime}_j$ by diagonal dominance of semidefinite matrices.

@@

We have seen that exact restricted isometry constants can be computed not only via total enumeration but also by solving a mixed integer semidefinite program. Because of this exactness result, this new problem could drastically
improve the bounds computed with the semidefinite relaxations proposed in \cite{Asp07} and \cite{Asp08}, but this came at the cost of a much higher complexity of computation. We could only solve these problems for relatively small
instances, but this could change with sufficient advances in the area of mixed integer semidefinite programming, especially with the addition of special cutting planes, which proved to be the main enhancement needed to make mixed
integer linear solvers as fast as they are today. Also special heuristics and branching rules could improve the performance of the branch-and-bound-algorithm and allow us to solve bigger instances of the proposed MISDP.

>>>>

Now let $\bar{X}$ be the $\ell \times \ell$ matrix with all rows and columns outside of $T$ removed. Then $\bar{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), 
as we only removed zero rows and columns of $\hat{X}$ and $(A\T A)\hat{X}$, which do not influence the trace or the positive semidefiniteness.

On the other hand we can lift any optimal solution $\check{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $z^{\prime}_i$ to one if and only if $i \in T$
and $X^{\prime}_{i_{j_1} i_{j_2}} = \check{X}_{j_1j_2}$ if and only if $i_{j_1}, i_{j_2} \in T$
and $X^{\prime}_{ij}=0$ otherwise. This solution will then be feasible for (\ref{MISDP}), as $\Tr(\check{X}) = 1$ and $\check{X}\in S_\ell$ imply $0 \leq \check{X}_{jj} = X^{\prime}_{i_j i_j} \leq 1$ and therefore 
$X^{\prime}_{ij} \leq z^{\prime}_j$ by diagonal dominance of semidefinite matrices.

@@

We have seen that exact restricted isometry constants can be computed not only via total enumeration but also by solving a mixed integer semidefinite program. Because of this exactness result, this new problem could drastically
improve the bounds computed with the semidefinite relaxations proposed in \cite{Asp07} and \cite{Asp08}, but this came at the cost of a much higher complexity of computation, at least in comparison to \cite{Asp07}. We could only 
solve these problems for relatively small instances, but this could change with sufficient advances in the area of mixed integer semidefinite programming, especially with the addition of special cutting planes, which proved to be 
the main enhancement needed to make mixed integer linear solvers as fast as they are today. Also special heuristics and branching rules could improve the performance of the branch-and-bound-algorithm and allow us to solve bigger 
instances of the proposed MISDP.
